---
title: "VariCat Cognitive Variables"
author: "Hui Xin Ng"
date: "2022-09-01"
output: html_document
---

```{r setup, include=FALSE}
# table that has the cognitive sub-domains with Number of studies and Median number of variables in that domain. 
# table of instruments that are used in more than 2 studies and how many studies they are used in.
library(readr)
library(tidyverse) 
library(ggplot2) 
```
 
# How many study sites are there for each cognitive subdomain?
```{r pressure, echo=FALSE}
data <- read_csv("gagebd_label_v2.csv")
data <- data[,c(4,5, 7, 1)] %>% replace_na(list(battery = 'None'))
names(data) <- c('cognitive_type',	'battery',	'instrumenttype','study')
df1 <- data %>% group_by(cognitive_type) %>% count(study)  %>% tally()
names(df1)[2] <- c("num_study")
 
# num of studies in each battery
data %>% replace_na(list(battery = 'None'))%>% group_by(battery) %>% count(study)   %>% tally() %>% filter(n>2)
 
```


```{r}
data <- data %>% mutate(instrumenttype =gsub("\\s*\\([^\\)]+\\)", "", instrumenttype))   %>% mutate(instrumenttype =gsub(",$", "", instrumenttype)) %>%   mutate(instrumenttype =str_to_title( instrumenttype))


# table
df1<- df1 %>% arrange(desc(num_study))
 data %>% mutate(instrumenttype =gsub("\\s*\\([^\\)]+\\)", "", instrumenttype))   %>% mutate(instrumenttype =gsub(",$", "", instrumenttype)) %>%   mutate(instrumenttype =str_to_title( instrumenttype))
 

```

```{r}  
other_instrument <- read_csv("other_instrument_v2.csv", show_col_types = FALSE)

other_instrument <- other_instrument  %>%
  mutate_if(is.character, str_trim)

colnames(other_instrument) <- c('record_id',	'cognitive_type', 	"instrumenttype", 	"instrumenttypeoth",	"study", 	'varname')

other_instrument <- other_instrument %>% mutate(instrumenttypeoth = case_when(instrumenttypeoth %in% c('D-KEFS','DKFES', 'Delis') ~ 'DKEFS', TRUE ~ instrumenttypeoth ))%>% 
  mutate(instrumenttypeoth = case_when(instrumenttypeoth %in% c('BAC-A: Brief Assessment of Cognition In Affective Disorders (BAC-A)','BAC-A:Brief Assessment of Cognition In Affective Disorders (BAC-A)' ) ~ 'BAC-A: brief assessment of cognition in affective disorders', TRUE ~ instrumenttypeoth )) %>%
  
  mutate(instrumenttypeoth = case_when(instrumenttypeoth %in% c('Memory for Intensions Screening Test (MIST)', 'Memory for Intentions Screening Test (MIST):'  ) ~ 'Memory for Intentions Screening Test (MIST)', TRUE ~ instrumenttypeoth )) %>% 
  
  mutate(instrumenttypeoth = case_when(instrumenttypeoth %in% c('FAB/ frontal assessment battery' ) ~ 'FAB-frontal assessment battery', TRUE ~ instrumenttypeoth ))
other_instrument <- other_instrument %>% 
  
  mutate(instrumenttypeoth = str_trim(instrumenttypeoth)) %>% mutate(instrumenttypeoth =gsub("\\-%$:$", "", other_instrument$instrumenttypeoth)) %>% mutate(instrumenttypeoth =gsub("Boston Naming Test", "", other_instrument$instrumenttypeoth))

other_instrument  %>%
  group_by(instrumenttypeoth, study)%>%count(instrumenttypeoth, study)   %>% replace_na(list(instrumenttypeoth = 'None'))  
```

 

CBBSX-GGZ (13), DOBI1 and DOBI2 used a modified version of the Auditory Verbal Learning Test

# What is the median, min and max number of variables for each cognitive domain?
```{r pressure, echo=FALSE} 
df2 <- data %>% group_by(cognitive_type) %>% count(study) %>% summarise( min = min(n, na.rm=TRUE), max = max(n, na.rm=TRUE), median_num_vars_per_cog_domain = median(n))
data %>% group_by(cognitive_type) %>% count(study) # check how many vars of a particular domain there are from a study 
```

# merge df1 and df2
```{r}
n_study_in_cog_domain <-merge(df1,df2, by=c("cognitive_type")) %>% arrange(desc(num_study))
 
write.csv(n_study_in_cog_domain, "table1.csv")
# Barplot

n_study_in_cog_domain #87 is from geri-sad, mostly DRS and MINI MMSE
```

```{r}
number_ticks <- function(n) {function(limits) pretty(limits, n)}

ggplot(n_study_in_cog_domain, aes(x=reorder(cognitive_type,num_study), fill=cognitive_type, y=num_study) ) + 
  geom_bar(stat = "identity",   show.legend = FALSE) +coord_flip()+theme_classic() +
  xlab("Cognitive Domain ") + ylab("Number of Studies") + scale_y_continuous(breaks=number_ticks(10)) # +  scale_fill_brewer(palette="Paired")
```
# How many instruments are used in more than 2 studies, and which cognitive domains do they belong to?
```{r}
more_than_2_times_instrument <- data %>% group_by(cognitive_type, instrumenttype)%>% count(study) %>% tally()  %>% arrange(desc(n), .by_group = TRUE,  ) %>% filter(., n>2, instrumenttype != 'Other' )   %>% mutate(instrumenttype =gsub("\\s*\\([^\\)]+\\)", "", instrumenttype))   %>% mutate(instrumenttype =gsub(",$", "", instrumenttype)) %>%   mutate(instrumenttype =str_to_title( instrumenttype))
write.csv(more_than_2_times_instrument, 'table2.csv')
# Executive Functioning	Trail Making Test Part B	10
more_than_2_times_instrument

# Procssing speed	Trail Making Test Part A 8
# Executive Functioning	Trail Making Test Part A	7		
 scaleFUN <- function(x) sprintf("%.0f", x)

freq_more_than_2_times_instrument <- more_than_2_times_instrument%>% ungroup()  %>% select(instrumenttype, n) %>% group_by(instrumenttype) %>% summarize(sum = sum(n))%>% filter(instrumenttype != "None")
ggplot(freq_more_than_2_times_instrument, aes(x=reorder(instrumenttype,  sum) #, fill = instrumenttype 
                                              , y=sum))+
  geom_bar(stat='identity',   show.legend = FALSE)+
  ylab("Number of Studies") +  coord_flip()+theme_classic()  +  xlab("Cognitive Domain")  +   scale_y_continuous(breaks=number_ticks(10)) # +  scale_color_brewer(palette="BrBG") 



freq_more_than_2_times_instrument <- more_than_2_times_instrument%>% ungroup() %>% filter(cognitive_type == "Executive Functioning") %>% select(instrumenttype, n)  %>% group_by(instrumenttype) %>% summarize(sum = sum(n)) %>% filter(instrumenttype != "None")
ggplot(freq_more_than_2_times_instrument, aes(x=reorder(instrumenttype,  sum)  , y=sum))+
  geom_bar(stat='identity',   show.legend = FALSE)+
  ylab("Number of Studies") +  coord_flip()+theme_classic()  +  xlab("Cognitive Domain")  +   scale_y_continuous(breaks=number_ticks(10))  +  scale_color_brewer(palette="BrBG")
# n refers to number of studies
```
### What domains have Other or NA as their instruments and have at least >2 studies using Other or NA?
```{r}
domains_withNAOther <- data %>% group_by(cognitive_type, instrumenttype)%>% count(study) %>% tally()  %>% arrange(desc(n), .by_group = TRUE,  ) %>% filter(., n>2, instrumenttype == 'Other'| is.na(instrumenttype)) %>% pull(cognitive_type)

data %>% group_by(cognitive_type, instrumenttype)%>% count(study)  %>% tally() %>% arrange(desc(n), .by_group = TRUE,  ) %>% filter(., n>2, instrumenttype == 'Other'| is.na(instrumenttype)) 
```



### Which domain has any Nones for instrument type? 
```{r}
data %>% group_by(cognitive_type, instrumenttype)%>% mutate(count = n_distinct(study))%>% filter(instrumenttype == "None")  #%>% filter(count >2)   
```
Global cognitive functioning variable has instrument "NA" in the four following studies:	Barcelona_Radua_wave2. CBBSX-GGZ, and Inflammaging		 

### MISC. 
# How many cognitive domains are there for each study site? 
```{r  }
data %>% group_by(study) %>% count(cognitive_type) %>% tally()
```

# How many studies was each instrument used in?
```{r}
data %>% group_by(study) %>% count(instrumenttype) %>% tally()
```

# How many instrument types did each study have?
```{r}
data %>% group_by(instrumenttype  ) %>% count(study) %>% tally() %>% filter(., n>2)
```

# Which domains had NA or Other as their instrument type?
```{r}
# other_instrument <- data %>% group_by(cognitive_type, instrumenttype)%>% mutate(count = n_distinct(study))%>% filter(instrumenttype == 'Other' ) %>% filter(cognitive_type %in% domains_withNAOther) %>% distinct()
# write.csv(other_instrument, 'other_instrument.csv')
# 
# other_instrument
```

