---
title: "Varicat Ontology Integrated Dataset and Inflammaging Dataset Ontology - Individual Pages"
author: "Hui Xin Ng"
date: "2023-03-03"
output: html_document
---


### Basic data cleaning functions and helper functions
```{r, echo=FALSE, message= FALSE}
x <- c("stringr", "ggraph", "igraph", "cli", "dplyr", "plyr", "tidyverse", "data.tree", "rlang", "networkD3", "htmltools", "stringr", "webshot")
lapply(x, FUN = function(X) {
    do.call("require", list(X)) 
}) 

# ==== Helper functions ====
ls <- c("type", "of")
## Construct the regular expression
pat <- paste0("\\b(", paste0(ls, collapse="|"), ")\\b")     

# Note that "Drug.class.type." refers to "current drug class" and and "Other.drug.type" refers to "Current Other"
clean_colnames <- function(ls) {
  ls <-gsub('[^[:alnum:]]', ' ', ls)

ls <- tolower(ls)
  ls <-gsub(pat, '', ls)
  ls <- trimws(ls)
    ls <- str_squish(ls)  
  return (ls)
} 

paste5 <- function(..., sep = " ", collapse = NULL, na.rm = F) {
  if (na.rm == F)
    paste(..., sep = sep, collapse = collapse)
  else
    if (na.rm == T) {
      paste.na <- function(x, sep) {
        x <- gsub("^\\s+|\\s+$", "", x)
        ret <- paste(na.omit(x), collapse = sep)
        is.na(ret) <- ret == ""
        return(ret)
      }
      df <- data.frame(..., stringsAsFactors = F)
      ret <- apply(df, 1, FUN = function(x) paste.na(x, sep))
      if (is.null(collapse))
        ret
      else {
        paste.na(ret, sep = collapse)
      }
    }
} 



```
### Read dataset (CHOOSE INFLAMMAGING OR INTEGRATED)
```{r}
integrated_gage_varicat <- read.csv("integrated_gage_varicat.csv", sep=",", header=TRUE) 


# Uncomment the two lines below to create the clusters for inflam dataset
inflam <- read.csv("inflammaging_varicat.csv", sep=",", header=TRUE)
integrated_gage_varicat <-  inflam 
```
### remove unnescessary info 
```{r}
colnames <- colnames(integrated_gage_varicat) 
 
# converting the result to dataframe
integrated_gage_varicat <- as.data.frame(integrated_gage_varicat) %>% mutate_all(., list(~na_if(.,""))) 
# NOTE: look up: regex and lapply
colnames<- colnames[lapply(colnames, function(x) length(grep("Please|Is.this|What|Was.|Do.we|How.many.|Variables.this|General.notes|Complete" ,x,value=FALSE))) == 0] # remove sublevels that are too
length(unique(colnames)) #75

# === classify each column name into levels 1, 2 vs. 3 ===  
 integrated_gage_varicat <-  integrated_gage_varicat[colnames] %>% select(-head(names(.), 2)) %>% mutate_all(., na_if, y = "") %>% distinct() 
 
integrated_gage_varicat <- as.data.frame(apply(integrated_gage_varicat,2, str_trim))   # remove trailing spaces
# subset(rev_df,duplicated(level3)) #check if there are duplicats in lvl 3 except for NAs

colnames(integrated_gage_varicat) <- clean_colnames(colnames(integrated_gage_varicat))
integrated_gage_varicat <- integrated_gage_varicat[,4:ncol(integrated_gage_varicat)]  
 
```
# Classify columns in the dataset into Main Domain, Sub-Domain and Sub-sub domain
```{r}
# get the col names of the level 1 categories 
l1Cols <- c("Type.of.clinical.characteristics.", "Clinical.Trial..Specific..type.", "Cognitive.type.", "Course.of.bipolar.illness.episodes.type.", 
           "Course.of.bipolar.illness.symptoms.type." ,   "Course.of.non.bipolar.psychiatric.illness.type.",  "Course.of.treatment.type." , "Current.illness.severity.type.", 
           "Current.pharmacological.treatment.type.",  "Type.of.demographics." ,   "Ecological.Momentary.Assessment.type.",  "Family.history.type.",
           "Lifetime.pharmacological.treatment.type.", "Type.of.Metadata.", "Miscellaneous.type." , "Other.non.pharmacological.treatment.type.", 
           "Physical.health.type.",  "Physiological.type.", "Positive.Psychological.Factors.type.", "Trauma.Stress..type.") 
 
l2Cols <- c("Comorbidities.type.", "Depressive.Episode.Course.type.", "Drug.class.type.", "ECT.type.", "First.episode.features.type.", "Functioning.type.", "General.Episodic.Features.type.", "Global.type.", "History.of.non.bipolar.diagnosis.type.", "Hypomanic.Episode.Course.type.", "Imaging.type.", "Lab.work.type.", "Lifetime.drug.class.type.", "Manic.Episode.Course.type.", "Medication.adherence.type.", "Mixed.Episode.Course.type.", "Physical.Abuse.type.", "Physical.Measurements.type.", "Psychotic.Episode.Course.type.", "Sexual.Abuse.type.", "Sleep.circadian.rhythms.type.", "Smoking.type.", "Substance.use.type.", "Type.of.current.diagnosis.", "Type.of.history.of.psych.symptom.variables.", "Type.of.hospitalization.variables.", "Type.of.suicidality.variables.")

l3Cols <- c("Anticonvulsants.Mood.Stabilizers.type.", "Antidepressant.type.", "Antipsychotic.type.", "Global.physical.comorbidity.type.", "Heart.type.","Li.Lithium..type.", "Lifetime.Anticonvulsants.Mood.Stabilizers.type.", "Lifetime.Antidepressant.type.", "Lifetime.Antipsychotic.type.", "Lifetime.Li.Lithium..type.", "Lifetime.Non.psychiatric.medication.type.", "Lifetime.Other.drug.type.", "Lifetime.Polypharmacy.drug.type.", "Lifetime.Sedative.Hypnotics.type.", "Lifetime.Stimulants.type.", "Non.psychiatric.medication.type.", "Other.drug.type.", "Polypharmacy.drug.type.", "Saliva..skin..hair.type.", "Sedative.Hypnotics.type.", "Stimulants.type.", "Vascular.type."   )

l1Cols <- clean_colnames(l1Cols)
l2Cols <- clean_colnames(l2Cols)
l3Cols <- clean_colnames(l3Cols)

### === 
l1_df <- cbind(integrated_gage_varicat[1], integrated_gage_varicat %>% select(l1Cols))
groupbyl1 <- cbind(l1_df[1], l2 = do.call(pmax, c(l1_df[-1], na.rm = TRUE))) #%>% distinct()   # collapse all level2 vars into one column (l2) and remove duplicate rows then group by the main domain  
l2_df <- cbind(integrated_gage_varicat[1], integrated_gage_varicat %>% select(l2Cols))
groupbyl2 <- cbind(l2_df[1], l3 = do.call(pmax, c(l2_df[-1], na.rm = TRUE))) #%>% distinct() 
colnames(integrated_gage_varicat)[colnames(integrated_gage_varicat) == "drug class"] ="current drug class"
```

```{r}
# add l2 and l3 columns to the original df. Now we collapse the multiple columns which have many NAs in their rows into one column based on whether they are l2 or l3
integrated_gage_varicat$l2 <- groupbyl1$l2
integrated_gage_varicat$l3 <-groupbyl2$l3
colnames(integrated_gage_varicat)[1] <- c("l1")  
# integrated_gage_varicat %>% arrange(l1)
df <- integrated_gage_varicat %>% relocate(l2, .after="l2") %>% relocate(l3, .after="l2") %>%  dplyr::select(l1,l2,l3)  %>% arrange(l1, l2, l3)
df <- df %>%  mutate(l2 = case_when(l2 == "Drug class" ~ "Current drug class" , 
                                    
                                    TRUE ~ l2))

df_n <- df %>% group_by(l2, l3) %>% dplyr::mutate(n = n())   %>% arrange(l1, l2, l3)%>% distinct() 
df_n  <- df_n  %>% dplyr::mutate(l2 = (ifelse(l2 == "Other", paste(l1, l2), l2)), 
                                 l3 = (ifelse(l3 == "Other", paste(l2, l3), l3)), 
                                 l1 = (ifelse(l1 == l2, paste(l1, "Domain"), l1))) %>% mutate(l3 = str_replace_all(l3, "/", "_"), 
                                                                           l2 = str_replace_all(l2, "/", "_"), 
                                                                           l1 = str_replace_all(l1, "/", "_"))   %>% mutate(l3 = str_replace_all(l3, " ", "_"), 
                                                                           l2 = str_replace_all(l2, " ", "_"), 
                                                                           l1 = str_replace_all(l1, " ", "_"))
df_n$l0 <- rep("Var", dim(df_n)[1])
df_n$pathString <- paste5(df_n$l0, 
                            df_n$l1, 
                            df_n$l2, 
                            df_n$l3,
                            sep = "/", na.rm=TRUE )  
grp <- df_n %>% group_by(l1) 
ls_dfs <- group_split(grp)   
```
### create function that takes in a data frame for visualization
```{r}
create_fn <- function(df_n)  {
  df_ontology <- as.Node(df_n) # create data.tree, 102 rows
  
  # https://rdrr.io/cran/data.tree/man/as.igraph.Node.html data.tree --> igraph keep vertex size info
  #print(df_ontology, "n")
  # convert data.tree to igraph obj
  df_g <- as.igraph(df_ontology, direction = "climb")  %>% delete_vertices("Var")  #%>%  delete_edges( c(1:10))  - can be used to delete vertices with string NA
   
  df_temp<- df_n %>% 
      arrange(l1, l2) %>% 
      mutate(l1_id = (factor(l1)))  # factorize
  
  ### Combine level 2 and level 3 variables into one single column 
  a <- cbind(df_n[2:3], name = do.call(pmax, c(df_n[4], na.rm = TRUE))) %>% na.omit()  %>% ungroup() %>% dplyr::select(l3, name)   
  colnames(a) <- c("name", "n") 
  b <- cbind(df_n[2:3], name = do.call(pmax, c(df_n[4], na.rm = TRUE)))%>% filter(is.na(l3)) %>% ungroup() %>% dplyr::select(l2, name)
  colnames(b) <- c("name", "n")
  df_combined_level_2_3 <- rbind(a, b)
  
  df_temp <- bind_rows(df_temp[c("l1", "l1_id")], df_temp[c( "l2" ,"l1_id")], df_temp[c("l3","l1_id")]) # create new df grouped by the domain
  membership_custom <- cbind(df_temp[2], name = do.call(pmax, c(df_temp[-2], na.rm = TRUE))) %>% distinct() %>% drop_na()  %>% arrange(name) # group l2 and l3 variables under l1
  df_d3 <- igraph_to_networkD3(df_g) # convert igraph obj to networkd3 obj
  
  membership_custom <-  merge(df_d3$nodes, membership_custom, on= "name") # merge node df and the new df grouped by domain to get desired color groupings
  df_d3$nodes <-  membership_custom[match(df_d3$nodes$name, membership_custom$name),] # order vars according to the data.tree obj's original order
  rnames <- rownames(df_d3$nodes)
  df_d3$nodes$idx <- as.integer(rownames(df_d3$nodes))
  
  join_df_combined_l2_l3_n <- full_join( df_d3$nodes, df_combined_level_2_3, on = "name")  %>% replace(is.na(.), 0)
  df_d3$nodes <- merge(df_d3$nodes, join_df_combined_l2_l3_n, by=c("name", "l1_id", "idx"), all.x = TRUE)# %>% dplyr::select()
  
  df_d3$nodes <- df_d3$nodes[as.integer(rnames), ] # order the rows by the original index number order
  
  add_n_if_sub_domain <- ifelse(!df_d3$nodes$name %in% df_d3$nodes$l1_id , paste0(df_d3$nodes$name, ": ", df_d3$nodes$n), df_d3$nodes$name)
  
  df_d3$nodes <- df_d3$nodes%>% mutate(name = add_n_if_sub_domain)
      
# make a grouping variable that will match to colours
 df_d3$nodes <- df_d3$nodes %>%
 mutate(color_group = case_when(
    name %in% c("Clinical_characteristics",
    "Cognitive",
    "Course_of_bipolar_illness-episodes",
    "Course_of_bipolar_illness-symptoms",
    "Course_of_non-bipolar_psychiatric_illness",
    "Current_illness_severity",
    "Current_pharmacological_treatment",
    "Demographics",
    "Ecological_Momentary_Assessment",
    "Family_history",
    "Lifetime_pharmacological_treatment",
    "Metadata",
    "Non-pharmacological_treatment",
    "Trauma_stress",
    "Physical_health",
    "Physiological",
    "Positive_psychological_factors") ~ "main",
     TRUE ~ paste0(l1_id, "_sub")
  ))
  
 
  return(df_d3)  
}

for (i in ls_dfs)    {
  print(unique(i$l1))
  print(sum(i$n))
}

  
``` 



```{r}
# these are the colors of each of the domain's clusters; this vector is not used but is listed here so we can see the colors lsited in "my_color" vector
c("#7FC07F" , "#BEAED4" ,"darkgreen" ,"#C90B20", "violet" ,
 "#F0027F" ,"#BF5B20" ,"#666666" ,"#1B0E77" ,"#D95F00" ,
"#7570B3" ,"#E7290A", "#66A61E" ,"#E6AA00" ,"#A6761D" ,
"#666666" ,"#A6CED3", "darkmagenta"  ,"#FB9A99",  "turquoise" )
 
   
my_color <- 'd3.scaleOrdinal()
  .domain([
  "Clinical_characteristics_sub", "Cognitive_sub",  "Course_of_bipolar_illness-episodes", "Course_of_bipolar_illness-symptoms_sub", "Course_of_non-bipolar_psychiatric_illness_sub", "Current_illness_severity_sub"   , "Current_pharmacological_treatment_sub", "Demographics_sub", "Ecological_Momentary_Assessment_sub"  , "Family_history_sub"  , "Lifetime_pharmacological_treatment_sub", "Metadata_sub"  ,  "Trauma_stress_sub"  ,  "Physical_health_sub", "Physiological_sub" ,  "Positive_psychological_factors_sub_sub", "Clinical_Trial-specific_sub_sub", "Course_of_treatment_sub",     "Non-pharmacological_treatment_sub", "Miscellaneous_Domain_sub", "main"
  ])
  .range([
     "#7FC07F",  
    "#BEAED4",
    "darkgreen",
    "#C90B20",
    "violet",
    "#F0027F",
    "#BF5B20",
    "#666666",
    "#1B0E77",
    "#D95F00",
    "#7570B3",
    "#E7290A",
    "#66A61E",
    "#E6AA00",
    "#A6761D",
    "#666666",
    "#A6CED3",
    "darkmagenta",
    "#FB9A99",
    "turquoise",
    "black" 
  ])' 
  

### === DEFINE FUNCTION TO CREATE FORCE NETWORK /D3 object === 
visualizeDomainNetwork <- function(df_n ,folder_and_filename_prefix) {
    viz_name <- unique(df_n$l1)
    df_d3 <- create_fn(df_n)



fn <- forceNetwork(Links = df_d3$links, Nodes = df_d3$nodes ,
             Source = 'source', Target = 'target', NodeID = 'name',  Group = 'color_group', # get NodeID and Group from inflam_d3$node
            opacity = 0.9,  zoom = TRUE,  linkDistance = 20,  #legend = TRUE,
             charge = -40,   # node repulsion   
            Nodesize  = "n" , colourScale = my_color, fontSize = 20, fontFamily = "Calibri", 
            )  
 

### === CHANGE FILE AND FOLDER NAME === 
filename = paste0(folder_and_filename_prefix, viz_name,".html")
jpegname = paste0(folder_and_filename_prefix, viz_name,".jpeg") 
 
 

clickjs <- "function(el, x) {
  var options = x.options; // Store the options passed from R

  // Select the SVG element
  var svg = d3.select(el).select('svg');

  // Select all nodes and links in the SVG
  var node = svg.selectAll('.node');
  var link = svg.selectAll('link');

  // Store the mouseout event listener for nodes
  var mouseout = d3.selectAll('.node').on('mouseout');

  // Function to calculate the node size based on options
  function nodeSize(d) {
    if (options.nodesize) {
      return eval(options.radiusCalculation);
    } else {
      return 6;
    }
  }

  // Add click event listener to all nodes
  d3.selectAll('.node').on('click', onclick);

  // Function to handle node click events
  function onclick(d) {
    if (d3.select(this).on('mouseout') == mouseout) {
      // If node is not clicked, assign mouseout_clicked event listener
      d3.select(this).on('mouseout', mouseout_clicked);
    } else {
      // If node is clicked, assign regular mouseout event listener
      d3.select(this).on('mouseout', mouseout);
    }
  }

  // Function to handle mouseout event for clicked nodes
  function mouseout_clicked(d) {
    // Reset opacity of nodes and links
    node.style('opacity', +options.opacity);
    link.style('opacity', +options.opacity);

    // Transition the node circle to its original size
    d3.select(this).select('circle').transition()
      .duration(750)
      .attr('r', function(d) { return nodeSize(d); });

    // Transition the node text to its original position and font size
    d3.select(this).select('text').transition()
      .duration(1250)
      .attr('x', 0)
      .style('font', options.fontSize + 'px ');
  }
  
 
  
}" 
 
htmlwidgets::onRender(fn, clickjs) %>%  saveNetwork(filename) 

# convert html file to jpeg
webshot(filename, jpegname, vwidth = 300, vheight = 300, zoom = 1.2)  
} 
 


```

### Save the individual clusters into separate html files 
```{r}

### ==== Specify the folder name, and filename prefix depending on whether it is inflammaging or integrated gage ====

#folder_and_filename_prefix = "integrated_gage/integrated_gage_"  #uncomment to create clusters for the integrated dataset
folder_and_filename_prefix ="inflammaging/inflammaging_" 

num_dfs = length(ls_dfs)-1

for(i in 1:num_dfs) {
 visualizeDomainNetwork(ls_dfs[[i]], folder_and_filename_prefix )
}   
 
```
 
 
 
### === ARCHIVED ===
### Decide on a color for each of the 20 Domains
```{r}
# Create networkD3 obj
df_d3 <- create_fn(df_n)  

# fn <- forceNetwork(Links = df_d3$links, Nodes = df_d3$nodes   ,colourScale  = my_color, 
#              Source = 'source', Target = 'target', NodeID = 'name',  Group = 'l1_id', # get NodeID and Group from inflam_d3$node
#             opacity = 0.9,  zoom = TRUE, legend = TRUE, linkDistance = 10,  
#              charge = -25,   # node repulsion   
#             Nodesize  = "n" , 
#          fontSize = 20, fontFamily = "Calibri" )
# fn  %>% saveNetwork("integrated_gage_test.html")
```

 